---
title: "Spatial data for: Drought may initiate western spruce budworm outbreaks, but multi-year periods of increased moisture availability promote widespread defoliation"
author: "Sarah J. Hart, Olivia Santiago, Joshua D. Carrell, and Thomas T. Veblen"
email: "sarah.hart@colostate.edu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
   officedown::rdocx_document:
     reference_docx: Template.docx
bibliography: references.bib
csl: "`r here::here('citationstyle.csl')`"
link-citations: true
urlcolor: blue
linkcolor: blue
citationcolor: blue
---

```{r setup,include=FALSE, results='hide'}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	progress = FALSE,
	cache = FALSE,
	dpi = 300,
  fig.align = 'center'
)

set.seed(513)
options(repos = c(CRAN = "http://cran.rstudio.com"))
options(timeout=60*60) # timeout downloads that last longer than 1 hour

### Import libraries
if (!require("remotes"))  install.packages("remotes")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here, # easy file structure
  rmarkdown, # 
  flextable, # tables
  officer, # word documents
  bookdown, # word documents
  tidyverse, # data manipulation
  ggplot2, # figures
  RColorBrewer, # color palettes
  patchwork, # combining figures
  sf, # spatial data
  terra, # raster data (new pkg)
  raster, # raster data (old pkg)
  ncdf4, #import ncdf files
  tmap, # easy plotting of maps
  tmaptools, # helper library for tmap
  grid, # useful for plotting inset map
  terrainr, # dem
  prism # prism data
) 

proj.proj <- 4629

# Set custom plotting theme
theme_new <- function(base_size = 9,base_family = "Helvetica"){
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.line.x = element_line(color="black", linewidth = 0.25),
      axis.line.y = element_line(color="black", linewidth = 0.25),
      axis.title = element_text(size = 9),
      axis.text = element_text(colour="black", size=8),
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      plot.background = element_rect(fill = NA, colour = NA),
      panel.border = element_rect(fill = NA, colour = NA),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(size = 9)
      
    )
}
theme_set(theme_new())

set_flextable_defaults(
  font.family="Times", 
  font.size=12,
  line_spacing=1,
  padding.bottom=1,
  padding.top=1,
  text.align='center')

# Set directory structure for project
dir.create(here("Data", "Spatial"), showWarnings = FALSE)
dir.create(here("Results"), showWarnings = FALSE)
dir.create(here("Results", "Figures"), showWarnings = FALSE)
```

# Overview

The following code will download nearly all publicly-available data used in this study. Where data is not publicly-available or easy to programmatically download, we provide more details about how to access the data.

# Tree ring data

## Host

Tree-ring data collected from Douglas fir trees for the purpose of reconstructing past western spruce budworm outbreaks are available on Dryad (<https://doi.org/10.5061/dryad.bnzs7h4n3>). In our analyses, we stored this data in the project subdirectory `~Data/TreeRing/Raw/Host/Veblen`.

## Non-host

We obtained all raw ring width data collected from ponderosa pine trees collected in the state of Colorado from the International Tree Ring Databank (<https://www.ncei.noaa.gov/products/paleoclimatology/tree-ring>). In our analyses, we stored this data in the project subdirectory `~Data/TreeRing/Raw/NonHost/ITRDB`.

# Spatial data

## Basic Geographic data

We obtained spatial data describing state boundaries from the US Census Bureau (<https://www.census.gov/geographies/mapping-files.html>).

```{r DL_States, echo=T}
# states
dir.create(here("Data", "Spatial", "States"), showWarnings = FALSE)
temp <- here("Data", "Spatial", "States", "cb_2018_us_state_20m.zip")
if(!file.exists(temp)){
  download.file("https://www2.census.gov/geo/tiger/GENZ2018/shp/cb_2018_us_state_20m.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "States"))
}
```

We obtained point data describing the location of Colorado cities and towns from the Colorado Department of Health and Environment Open Data Portal (<https://data-cdphe.opendata.arcgis.com/>).

```{r DL_Cites, echo=T}
# Colorado cities
dir.create(here("Data", "Spatial", "Cities"), showWarnings = FALSE)
temp <- here("Data", "Spatial", "Cities", "Colorado_City_Point_Locations.zip")
if(!file.exists(temp)){
  download.file("https://hub.arcgis.com/api/v3/datasets/b20b5c04576145be9428b4b4f28490c2_0/downloads/data?format=shp&spatialRefId=4269&where=1%3D1", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "Cities"))
}
```

### Ecoregions

We obtained Level III Ecoregion data for the United States from the US Environmental Protection Agency (<https://www.epa.gov/eco-research/ecoregions-north-america>).

```{r DL_EcoRegions, echo=T}
#EcoRegions
dir.create(here("Data","Spatial", "EcoRegions"), showWarnings = FALSE)
temp <- here("Data", "Spatial", "EcoRegions","us_eco_l3.zip")
if(!file.exists(temp)){
  download.file("https://dmap-prod-oms-edc.s3.us-east-1.amazonaws.com/ORD/Ecoregions/us/us_eco_l3.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "EcoRegions"))
}
```

### Digital Elevation Model

We obtained a 250 m resolution Digital Elevation Model (DEM) for the study area from the National Elevation Dataset (NED) using the *terrainr* package [@terrainr].

```{r DEM, echo=T}
if(!file.exists(here("Data", "Spatial", "DEM", "DEM250.tif"))){
  # Southern Rockies
  srm <- st_read(here("Data", "Spatial", "EcoRegions", "us_eco_l3.shp")) %>% filter(US_L3NAME == 'Southern Rockies') %>% st_transform(crs=5071) %>% st_buffer(20000)
  dir.create(here("Data", "Spatial","DEM"), showWarnings = FALSE)

 # 250 m 
  dem <- get_tiles(data=srm, output_prefix = here("Data", "Spatial", "DEM", "DEM250"), services="elevation",   resolution=250)
  dem.rast <- rast(dem$elevation)
  writeRaster(dem.rast, here("Data", "Spatial", "DEM", "DEM250.tif"), overwrite=T)
}
```

## Douglas fir distribution

To characterize the distribution of Douglas fir across the study area, we acquired a 240 x 240 m raster describing tree presence intensity for Douglas fir in ca. 2002 from the USFS's Individual Tree Species Parameter Maps [@ellenwood2015]. Intensity was calculated by summing the number of a possible sixty four 30-meter pixels within a 240-meter pixel with presence of Douglas fir. The 2002 data is not available online, but can be obtained from the USFS (for more details see here <https://www.fs.usda.gov/science-technology/data-tools-products/fhp-mapping-reporting/individual-tree-species-parameter-maps>). In our analyses, we obtained this data in the form of an ArcInfo Gridstore and stored this data in the project subdirectory `~Data/Spatial/ITSPM/f202`.

## Disturbance data

We acquired polygon data describing the extent of burned areas for the 1984-2023 period from the Monitoring Trends in Burn Severity Project [-@mtbsprojectMTBSDataAccess2022] ([mtbs.gov](mtbs.gov)).

```{r DL_MTBS, echo=T}
# MTBS 
dir.create(here("Data", "Spatial", "MTBS"), showWarnings = FALSE)
temp <- here("Data", "Spatial", "MTBS", "mtbs_perimeter_data.zip")
if(!file.exists(temp)){
  download.file("https://edcintl.cr.usgs.gov/downloads/sciweb1/shared/MTBS_Fire/data/composite_data/burned_area_extent_shapefile/mtbs_perimeter_data.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "MTBS"))
}

```

### Aerial Detection Survey Data

To characterize recent history of WSB outbreaks across the study area, we acquired Aerial Detection Survey data for Region 2 from the USFS [-@usfsanditspartners2020USDAForestService] (<https://www.fs.usda.gov/science-technology/data-tools-products/fhp-mapping-reporting/detection-surveys>).

```{r DL_ADS, echo=T}
dir.create(here("Data", "Spatial", "ADS"), showWarnings = FALSE)
temp <- here("Data", "Spatial", "ADS", "CONUS_Region2_AllYears.gdb.zip")
if(!file.exists(temp)){
  download.file("https://www.fs.usda.gov/foresthealth/docs/IDS_Data_for_Download/CONUS_Region2_AllYears.gdb.zip", temp)
  unzip(temp, exdir=here("Data", "Spatial", "ADS"))
}
```

## Climate data

### North American Drought Atlas

We obtained multi-century (i.e., AD 1650-2005) records of the self-calibrating Palmer Drought Severity Index (SC-PDSI) from the North American Drought Atlas (NADA), which provides tree-ring based reconstructions of June-August SC-PDSI on a 0.5° resolution grid [@cook2010].

```{r DL_NADA, echo=T}
if(!dir.exists(here("Data", "Spatial", "NADA"))){
  dir.create(here("Data", "Spatial", "NADA"), showWarnings = FALSE)
  temp <- here("Data", "Spatial", "NADA", "nada_hd2_cl.nc")
  download.file("https://www.ncei.noaa.gov/pub/data/paleo/drought/LBDA2010/nada_hd2_cl.nc", temp)
}
```

### PRISM

To characterize the study area's climate, we obtained monthly precipitation totals, minimum temperatures, and maximum temperature 30-year normals from PRISM [@prismclimategroup2021] using the *prism* package [@prism]. We also obtained gridded June-August SC-PDSI data from the West Wide Drought Tracker [@abatzoglouWestWideDrought2017].

```{r DL_PRISM, echo=T}
if(!dir.exists(here("Data", "Spatial", "PRISM"))){
  dir.create(here("Data", "Spatial", "PRISM"), showWarnings = FALSE)
  prism_set_dl_dir(here("Data", "Spatial", "PRISM"))
  
  # Climate normals
  get_prism_normals("ppt", "800m", annual = TRUE, keepZip = FALSE)
  get_prism_normals("tmin", "800m", mon=1, keepZip = FALSE)
  get_prism_normals("tmax", "800m", mon=7, keepZip = FALSE)
  
  # Monthly climate data
  temp <- here("Data", "Spatial", "PRISM", "scpdsi_6_PRISM.nc")
  download.file("http://www.wrcc.dri.edu/wwdt/data/PRISM/scpdsi/scpdsi_6_PRISM.nc", temp)

  temp <- here("Data", "Spatial", "PRISM", "scpdsi_7_PRISM.nc")
  download.file("http://www.wrcc.dri.edu/wwdt/data/PRISM/scpdsi/scpdsi_7_PRISM.nc", temp)

  temp <- here("Data", "Spatial", "PRISM", "scpdsi_8_PRISM.nc")
  download.file("http://www.wrcc.dri.edu/wwdt/data/PRISM/scpdsi/scpdsi_8_PRISM.nc", temp)
}
```

### Climatic Water Balance

We also downloaded actual evaportranspiration (AET) and climatic water deficit (CWD) normals (1981-2010) from Rodman et al. [-@rodman2020DataTraitbasedApproach; -@rodman2020TraitbasedApproachAssessing]. Data is publicly available on Dryad (<https://datadryad.org/stash/dataset/doi:10.5061/dryad.cz8w9gj1b>). For our analyses, we downloaded the zipped data file and stored it in the project subdirectory `~Data/Spatial/Rodman`.

# References

::: {#refs}
:::