---
title: "Preprocessing of Douglas-fir chronologies for: Drought may initiate western spruce budworm outbreaks, but multi-year periods of increased moisture availability promote widespread defoliation"
author: "Sarah J. Hart, Olivia Santiago, Joshua D. Carrell, and Thomas T. Veblen"
email: "sarah.hart@colostate.edu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
   officedown::rdocx_document:
     reference_docx: Template.docx
bibliography: references.bib
csl: "`r here::here('citationstyle.csl')`"
link-citations: true
urlcolor: blue
linkcolor: blue
citationcolor: blue
---

```{r setup,include=FALSE, results='hide'}
knitr::opts_chunk$set(
  echo = TRUE,
	message = FALSE,
	warning = FALSE,
	progress = FALSE,
	cache = FALSE,
	dpi = 300,
  fig.align = 'center'
)

set.seed(513)
options(repos = c(CRAN = "http://cran.rstudio.com"))
options(timeout=60*30) # timeout downloads that last longer than 30 minutes

### Import libraries
if (!require("remotes"))  install.packages("remotes")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here, # easy file structure
  rmarkdown, # markdown
  flextable, # table creation
  officer, # word document creation
  officedown, # word document creation
  tidyverse, # data manipulation
  dplR, # basic dendrochronology
  dfoliatR, # dendroecology 
  zoo # time series
) 

proj.proj <- 4629

source(here("Code", "TreeRingfuncs.R"))

### Set table defaults 
set_flextable_defaults(
  font.family="Times", 
  font.size=12,
  line_spacing=1,
  padding.bottom=1,
  padding.top=1,
  text.align='center')

# Set directory structure for project
dir.create(here("Data"), showWarnings = FALSE)
dir.create(here("Results"), showWarnings = FALSE)
dir.create(here("Results", "Figures"), showWarnings = FALSE)
dir.create(here("Data", "TreeRing", "Processed"), showWarnings = FALSE)
dir.create(here("Data", "TreeRing", "Processed", "Host"), showWarnings = FALSE)
```

# Overview

Here we re-evaluated the crossdating of the host chronologies to ensure that the series were correctly dated. We used the dplR package [@bunn2008; @bunn2024DplRDendrochronologyProgram] to calculate the correlation between each series and a main chronology built from all other series at each site. Using the *corr.rwl.seg()* function, we calculated correlation coefficients in 50-year segments that overlapped by 25 years and then removed series that were poorly correlated with the main chronology.

### Boulder County 18

We retained most series from the Boulder County 18 site, but removed two series because they were poorly correlated with the other series (Fig. \@ref(fig:B18fig); Table \@ref(tab:B18tab)).

```{r B18dat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "B18.rwl"))
```

```{r B18fig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Boulder County 18 site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r B18tab, tab.id="B18tab", tab.cap="Series statistics and cross-dating notes for the Boulder County 18 site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))

colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")
summary.tab[summary.tab$Series %in% c("300", "276"), "Notes"] <- "Removed - poorly correlated with the other series"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r B18write, results=F}
rwlj <- rwlj[,!(colnames(rwlj) %in% c("300", "276"))]
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "B18.rwl"), format="tucson")
```

### Boulder County 19

We retained most series from the Boulder County 19 site, but removed three series because they were poorly correlated with the other series (Fig. \@ref(fig:B19fig); Table \@ref(tab:B19tab)).

```{r B19dat, results=F, echo=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "B19.rwl"))
```

```{r B19fig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Boulder County 19 site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r B19tab, tab.id="B19tab", tab.cap="Series statistics and cross-dating notes for the Boulder County 19 site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))

colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")
summary.tab[summary.tab$Series %in% c("458", "497", "443"), "Notes"] <- "Removed - poorly correlated with the other series"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1) %>% paginate()
```

```{r B19write, results=F}
rwlj <- rwlj[,!(colnames(rwlj) %in% c("458", "497", "443"))]
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "B19.rwl"), format="tucson")
```

### Lost Junction

We retained all series from the Lost Junction site (Fig. \@ref(fig:LJfig); Table \@ref(tab:LJtab)).

```{r LJdat, results=F, echo=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "LJ.rwl"))
```

```{r LJfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Lost Junction site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r LJtab, tab.id="LJtab", tab.cap="Series statistics and cross-dating notes for the Lost Junction site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r LJwrite, results=F}
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "LJ.rwl"), format="tucson")
```

### Sugarloaf Saddle

We retained all series from the Sugarloaf Saddle site (Fig. \@ref(fig:SSfig); Table \@ref(tab:SStab)).

```{r SSdat, results=F, echo=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "SS.rwl"))
```

```{r SSfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Sugarloaf Saddle site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r SStab, tab.id="SStab", tab.cap="Series statistics and cross-dating notes for the Sugarloaf Saddle site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r SSwrite, results=F}
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "SS.rwl"), format="tucson")
```

### Estes Park

We retained all series from the Estes Park site, but trimmed one series because it was poorly correlated at the beginning of the record (Fig. \@ref(fig:EPfig); Table \@ref(tab:EPtab)).

```{r EPdat, results=F, echo=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "EP.rwl"))
```

```{r EPfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series the Estes Park site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwlcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r EPtab, tab.id="EPtab", tab.cap="Series statistics and cross-dating notes for the Estes Park site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[summary.tab$Series=="632", "Notes"] <- "Years prior to 1750 were removed due to poor correlation"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r EPwrite, results=F}
rwlj <- clipfun(rwlz=rwlj, series.name="632", clip.year=1750)
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "EP.rwl"), format="tucson")
```

### Frankenberger Point

We retained all series from the Frankenberger Point site, but trimmed two series because they were poorly correlated at the beginning of the record (Fig. \@ref(fig:FPfig); Table \@ref(tab:FPtab)).

```{r FPdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen",  "FP.rwl"))
```

```{r FPfig, fig.width=4.5, fig.height=4.5, fig.cap="The Spearman's correlations between each tree-ring series and a main chronology built from all other series from the Frankenberger Point site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r FPtab, tab.id="FPtab", tab.cap="Series statistics and cross-dating notes for the Frankenberger Point site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[summary.tab$Series %in% c("CO95788", "CO95793"), "Notes"] <- "Removed years prior to 1700 due to poor correlation"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r FPwrite, results=F}
rwlj <- clipfun(rwlz=rwlj, series.name="CO95788", clip.year=1700)
rwlj <- clipfun(rwlz=rwlj, series.name="CO95793", clip.year=1700)
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "FP.rwl"), format="tucson")
```

### South Hollowell

We retained all series from the South Hollowell site (Fig. \@ref(fig:SHfig); Table \@ref(tab:SHtab)).

```{r SHdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "SH.rwl"))
```

```{r SHfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the South Hollowell site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r SHtab, tab.id="SHtab", tab.cap="Series statistics and cross-dating notes for the South Hollowell site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))

colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r SHwrite, results=F}
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "SH.rwl"), format="tucson")
```

### West Rd. 211

We retained all series from the West Rd. 211 site (Fig. \@ref(fig:WRfig); Table \@ref(tab:WRtab)).

```{r WRdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen",  "WR.rwl"))
```

```{r WRfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the West Rd. 211 site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r WRtab, tab.id="WRtab", tab.cap="Series statistics and cross-dating notes for the West Rd. 211 site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))

colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r WRwrite, results=F}
### Write to file
write.rwl(rwlj,fname=here("Data", "TreeRing", "Processed", "Host", "WR.rwl"), format="tucson")
```

### Lost Jeep

We retained all series from the Lost Jeep site (Fig. \@ref(fig:JPfig); Table \@ref(tab:JPtab)).

```{r JPdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "JP.rwl"))
```

```{r JPfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Lost Jeep site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r JPtab, tab.id="JPtab", tab.cap="Series statistics and cross-dating notes for the Lost Jeep site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[summary.tab$Series %in% c("390"), "Notes"] <- "Removed - poorly correlated with the other series"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r JPwrite, results=F}
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "JP.rwl"), format="tucson")
```

### Wigwam 1

We retained all series from the Wigwam 1 site (Fig. \@ref(fig:WWfig); Table \@ref(tab:WWtab)).

```{r WWdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "WW.rwl"))
```

```{r WWfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Wigwam 1 site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r WWtab, tab.id="WWtab", tab.cap="Series statistics and cross-dating notes for the Wigwam 1 site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r WWwrite, results=F}
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "WW.rwl"), format="tucson")
```

### South Rd. 30

We retained all series from the South Rd. 30 site (Fig. \@ref(fig:SRfig); Table \@ref(tab:SRtab)).

```{r SRdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "SR.rwl"))
```

```{r SRfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the the South Rd. 30 site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r SRtab, tab.id="SRtab", tab.cap="Series statistics and cross-dating notes for the South Rd. 30 site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[summary.tab$Series %in% c("300", "276"), "Notes"] <- "Removed - poorly correlated with the other series"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r SRwrite, results=F}
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "SR.rwl"), format="tucson")
```

### Summerland Park

We retained most series from the Summerland Park site, but removed two series because they were poorly correlated with the other series (Fig. \@ref(fig:SPfig); Table \@ref(tab:SPtab)).

```{r SPdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "SP.rwl"))
```

```{r SPfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Summerland Park site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r SPtab, tab.id="SPtab", tab.cap="Series statistics and cross-dating notes for the Summerland Park site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))

colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")
summary.tab[summary.tab$Series %in% c("NI1565", "NI1537"), "Notes"] <- "Removed - poorly correlated with the other series"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r SPwrite, results=F}
rwlj <- rwlj[,!(colnames(rwlj) %in% c("NI1565", "NI1537"))]
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "SP.rwl"), format="tucson")
```

### North Inlet

We retained most series from the North Inlet site, but removed ones series because it was poorly correlated with the other series (Fig. \@ref(fig:NIfig); Table \@ref(tab:NItab)).

```{r NIdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen",  "NI.rwl"))
```

```{r NIfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the North Inlet site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r NItab, tab.id="NItab", tab.cap="Series statistics and cross-dating notes for the North Inlet site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[summary.tab$Series %in% c("1120"), "Notes"] <- "Removed - poorly correlated with the other series"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r NIwrite, results=F}
rwlj <- rwlj[,!(colnames(rwlj) %in% c("1120"))]
### Write to file
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "NI.rwl"), format="tucson")
```

### Three Island Lake Trail

We retained all series from the Three Island Lake Trail (Fig. \@ref(fig:TIfig); Table \@ref(tab:TItab)).

```{r TIdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "TI.rwl"))
```

```{r TIfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Three Island Lake Trail site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r TItab, tab.id="TItab", tab.cap="Series statistics and cross-dating notes for the Three Island Lake Trail site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))

colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r TIwrite, results=F}
write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "TI.rwl"), format="tucson")
```

### Wild Basin

We retained all series from the Wild Basin site, but trimmed one series to the period following 1800 because of poor correlation in the early part of the record (Fig. \@ref(fig:WBfig); Table \@ref(tab:WBtab)).

```{r WBdat, results=F}
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Host", "Veblen", "WB.rwl"))
```

```{r WBfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series for the Wild Basin site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwljcor <- corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1)
```

```{r WBtab, tab.id="WBtab", tab.cap="Series statistics and cross-dating notes for the Wild Basin site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlj)
summary.tab$rho <- round(corr.rwl.seg(rwlj, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab$notes <- "Retained"

summary.tab <- summary.tab %>% select(series, first, last, rho, notes) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation", "Notes")

summary.tab[summary.tab$Series %in% c("WB1367"), "Notes"] <- "Trimmed series to remove all years prior to 1800 due to poor correlation"

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1) %>% paginate
```

```{r WBwrite, results=F}
rwlj <- clipfun(rwlz=rwlj, series.name="WB1367", clip.year=1800)

write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Host", "WB.rwl"), format="tucson")
```

```{r, results=F}
keymet.host <- read.csv(here("Data", "TreeRing", "Raw", "Host", "Veblen", "SiteInfo-WSB.csv"))
keymet.host$nseries=NA
keymet.host$FirstYear=NA
keymet.host$LastYear=NA
keymet.host$interrbar=NA
keymet.host$interrbar.sd=NA
keymet.host$ar1bar=NA
keymet.host$ar1bar.sd=NA


### Calculate basic chronology statistics
for(j in 1:nrow(keymet.host)){
  series <- as.character(keymet.host[j, "SeriesCode"])
  site <- keymet.host[j, "SiteName"]
  series.site <- gsub( " ", "-",paste(series, site))
  rwlj <- read.rwl(here("Data", "TreeRing", "Processed", "Host", paste0(series, ".rwl")))
  rwl.reportj <- rwl.report(rwlj)
  keymet.host[keymet.host$`SeriesCode`==series, "nseries"] <- rwl.reportj$nSeries # calculate the number of series
  keymet.host[keymet.host$`SeriesCode`==series, "FirstYear"] <- rwl.reportj$firstYear# calculate the number of series
  keymet.host[keymet.host$`SeriesCode`==series, "LastYear"] <- rwl.reportj$lastYear# calculate the number of series
  keymet.host[keymet.host$`SeriesCode`==series, "interrbar"] <- rwl.reportj$meanInterSeriesCor # calculate mean interseries correlation
  keymet.host[keymet.host$`SeriesCode`==series, "interrbar.sd"] <- rwl.reportj$sdInterSeriesCor# calculate sd interseries correlation
  keymet.host[keymet.host$`SeriesCode`==series, "ar1bar"] <- rwl.reportj$meanAR1 # calculate mean first order autocorrelation
  keymet.host[keymet.host$`SeriesCode`==series, "ar1bar.sd"] <- rwl.reportj$sdAR1 # calculate sd first order autocorrelation
}

write.csv(keymet.host, here("Data", "TreeRing", "Processed", "Host", "Host-Metadata.csv"), row.names=F)

```

# References

::: {#refs}
:::
