---
title: "Preprocessing of ponderosa pine chronologies for: Drought may initiate western spruce budworm outbreaks, but multi-year periods of increased moisture availability promote widespread defoliation"
author: "Sarah J. Hart, Olivia Santiago, Josh Carrell, and Thomas T. Veblen"
email: "sarah.hart@colostate.edu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
   officedown::rdocx_document:
     reference_docx: Template.docx
bibliography: references.bib
csl: "`r here::here('citationstyle.csl')`"
link-citations: true
urlcolor: blue
linkcolor: blue
citationcolor: blue
---

```{r setup,include=FALSE, results='hide'}
knitr::opts_chunk$set(
  echo = TRUE,
	message = FALSE,
	warning = FALSE,
	progress = FALSE,
	cache = FALSE,
	dpi = 300,
  fig.align = 'center'
)

set.seed(513)
options(repos = c(CRAN = "http://cran.rstudio.com"))
options(timeout=60*30) # timeout downloads that last longer than 30 minutes

### Import libraries
if (!require("remotes"))  install.packages("remotes")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here, # easy file structure
  rmarkdown, # markdown
  flextable, # table creation
  officer, # word document creation
  officedown, # word document creation
  tidyverse, # data manipulation
  dplR, # basic dendrochronology
  dfoliatR, # dendroecology 
  zoo # time series
) 

proj.proj <- 4629

source(here("Code", "TreeRingfuncs.R"))

### Set table defaults 
set_flextable_defaults(
  font.family="Times", 
  font.size=12,
  line_spacing=1,
  padding.bottom=1,
  padding.top=1,
  text.align='center')

# Set directory structure for project
dir.create(here("Data", "TreeRing", "Processed"), showWarnings = FALSE)
dir.create(here("Data", "TreeRing", "Processed", "NonHost"), showWarnings = FALSE)
```

# Overview

Here we compile metadata for ponderosa pine chronologies used as a non-host control series in our analyses. We also combine data from the same site that were collected in several collection efforts.

# Compile metadata

Additionally, we used the *dplR* package [@bunn2008; @bunn2024DplRDendrochronologyProgram] to calculate the number of series, chronology length, mean and standard deviation of the interseries correlation, and the mean and standard deviation of the first order autocorrelation coefficient.

## International Tree Ring Database

We downloaded all available ponderosa pine chronologies collected in the state of Colorado from the International Tree Ring Databank (ITRDB).

```{r KeyMetaITDRB, results=F, echo=F}

#### ITDRB ####
# list files downloaded from the ITDRB
x <- list.files(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa"), pattern="-noaa.rwl", full.names=T)

# set up empty data frame to hold results
N <- length(x)
keymet <- data.frame(SeriesCode= character(N), SiteName= character(N), Contributors= character(N), Lat=NA, Lon=NA, Elev=numeric(N), nseries=NA, FirstYear=NA, LastYear=NA, interrbar=NA, interrbar.sd=NA, ar1bar=NA, ar1bar.sd=NA)
  
# iterate through files and extract metadata
for(j in 1:N){
  filej <- x[[j]]
  keymet[j, "SeriesCode"]<- strsplit(readLines(con=filej, n=1), split="ITRDB ")[[1]][2]
  keymet[j, "SiteName"]<- strsplit(readLines(con=filej, n=1), split=" - ")[[1]][2]
  keymet[j, "Contributors"]<- strsplit(readLines(con=filej, n=35)[35], split="Investigators: ")[[1]][2]
  keymet[j, "Lat"]<- strsplit(readLines(con=filej, n=55)[55], split="# Northernmost_Latitude: ")[[1]][2]
  keymet[j, "Lon"]<- strsplit(readLines(con=filej, n=57)[57], split="# Easternmost_Longitude: ")[[1]][2]
  keymet[j, "Elev"]<- strsplit(readLines(con=filej, n=59)[59], split="# Elevation: ")[[1]][2]
}

### Calculate basic chronology statistics
for(j in 1:nrow(keymet)){
  series <- as.character(keymet[j, "SeriesCode"])
  site <- keymet[j, "SiteName"]
  series.site <- gsub( " ", "-",paste(series, site))

  filej <- list.files(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa"),pattern=paste0(series, ".rwl"), full.names=T, ignore.case=T)
  
  skip = FALSE
  
  tryCatch(rwlj  <- read.rwl(filej), error = function(e) { skip = TRUE})
   if(skip==TRUE) {next} else{
    rwl.reportj <- rwl.report(rwlj)
    rwl.summaryj <- summary.rwl(rwlj)
     keymet[keymet$`SeriesCode`==series, "nseries"] <- rwl.reportj$nSeries # calculate the number of series
     keymet[keymet$`SeriesCode`==series, "FirstYear"] <- rwl.reportj$firstYear# calculate the first year 
     keymet[keymet$`SeriesCode`==series, "LastYear"] <- rwl.reportj$lastYear# calculate the last year
  
    keymet[keymet$`SeriesCode`==series, "interrbar"] <- rwl.reportj$meanInterSeriesCor # calculate mean interseries correlation
    keymet[keymet$`SeriesCode`==series, "interrbar.sd"] <- rwl.reportj$sdInterSeriesCor# calculate sd interseries correlation
    keymet[keymet$`SeriesCode`==series, "ar1bar"] <- rwl.reportj$meanAR1 # calculate mean first order autocorrelation
    keymet[keymet$`SeriesCode`==series, "ar1bar.sd"] <- rwl.reportj$sdAR1 # calculate sd first order autocorrelation
    
    #### Write to file
    write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Nonhost", paste0(series, ".rwl")), format="tucson")
  }
}

```

## Veblen

We also acquired ponderosa pine chronologies initially collected by Veblen et al. [-@veblen2000] and subsequently recollected by Smith and Veblen in 2007.

```{r KeyMetaVeblen, results=F, echo=F}
#### VEBLEN LAB ####
DRI.rwl <- here("Data", "TreeRing", "Raw", "Nonhost", "Veblen",'DRI.rwl')
DRIre.rwl <- here("Data", "TreeRing", "Raw", "Nonhost", "Veblen", 'DRIre.rwl')
ELDOre.rwl <- here("Data", "TreeRing", "Raw", "Nonhost", "Veblen", 'ELDOre.rwl')
JEFFre.rwl <- here("Data", "TreeRing", "Raw", "Nonhost", "Veblen", 'JEFFre.rwl')
VANBre.rwl <- here("Data", "TreeRing", "Raw", "Nonhost", "Veblen", 'VANBre.rwl')
veblen.series <- list(DRI.rwl, DRIre.rwl, ELDOre.rwl, JEFFre.rwl, VANBre.rwl)
names(veblen.series) <- c("DRI", "DRIre", "ELDOre", "JEFFre", "VANBre")

keymet.veblen <- read.csv(here("Data", "TreeRing", "Raw", "Nonhost", "Veblen", "SiteInfo.csv"))
keymet.veblen$nseries=NA
keymet.veblen$FirstYear=NA
keymet.veblen$LastYear=NA
keymet.veblen$interrbar=NA
keymet.veblen$interrbar.sd=NA
keymet.veblen$ar1bar=NA
keymet.veblen$ar1bar.sd=NA


### Calculate basic chronology statistics
for(j in 1:nrow(keymet.veblen)){
  series <- as.character(keymet.veblen[j, "SeriesCode"])
  site <- keymet.veblen[j, "SiteName"]
  series.site <- gsub( " ", "-",paste(series, site))
  rwlj <- read.rwl(veblen.series[[series]])
  rwl.reportj <- rwl.report(rwlj)
  keymet.veblen[keymet.veblen$`SeriesCode`==series, "nseries"] <- rwl.reportj$nSeries # calculate the number of series
  keymet.veblen[keymet.veblen$`SeriesCode`==series, "FirstYear"] <- rwl.reportj$firstYear# calculate the number of series
  keymet.veblen[keymet.veblen$`SeriesCode`==series, "LastYear"] <- rwl.reportj$lastYear# calculate the number of series
  keymet.veblen[keymet.veblen$`SeriesCode`==series, "interrbar"] <- rwl.reportj$meanInterSeriesCor # calculate mean interseries correlation
  keymet.veblen[keymet.veblen$`SeriesCode`==series, "interrbar.sd"] <- rwl.reportj$sdInterSeriesCor# calculate sd interseries correlation
  keymet.veblen[keymet.veblen$`SeriesCode`==series, "ar1bar"] <- rwl.reportj$meanAR1 # calculate mean first order autocorrelation
  keymet.veblen[keymet.veblen$`SeriesCode`==series, "ar1bar.sd"] <- rwl.reportj$sdAR1 # calculate sd first order autocorrelation
  
    #### Write to file
    write.rwl(rwlj, fname=here("Data", "TreeRing", "Processed", "Nonhost", paste0(series, ".rwl")), format="tucson")

}

keymet <- rbind(keymet, keymet.veblen)
```

# Combine recollections with original data

We computed the correlation between each tree-ring series using the *corr.rwl.seg()* function in the *dplR* package [@bunn2008; @bunn2024DplRDendrochronologyProgram].

## Deer Mountain

We combined ponderosa pine data collected by Graybill [-@graybill2002NOAAWDSPaleoclimatology] and Woodhouse et al. [-@woodhouse2006NOAAWDSPaleoclimatologyc].

```{r DMdat, results=F, echo=F}
## CO532
filej <- list.files(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa"), pattern=paste0("CO532", ".rwl"), full.names=T, ignore.case=T)
rwlj <-read.rwl(filej)

# CO598
filej <- list.files(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa"), pattern=paste0("CO598", ".rwl"), full.names=T, ignore.case=T)
rwlj1 <-read.rwl(filej)
x <- which(colnames(rwlj1) %in% colnames(rwlj))
rwlj1 <- rwlj1[,-x]

### Combine
rwlcombo <- combine.rwl(rwlj, rwlj1)
```

```{r, DMfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Deer Mountain. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwlcombocor <- corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1)
```

```{r DMtab, tab.id="DMtab", tab.cap="Series statistics and cross-dating notes for the Deer Mountain site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlcombo)
summary.tab$rho <- round(corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab <- summary.tab %>% dplyr::select(series, first, last, rho) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation")
summary.tab$Update <- ifelse(summary.tab$Series %in% colnames(rwlj1), "Yes", "No")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r DMMeta, results=F, echo=F}
## Compile metadata ##
keymet.combo <- keymet %>% filter(SeriesCode=="CO532") 
keymet.combo$SeriesCode <- "DMcombo"
keymet.combo$Contributors <- "Graybill, D.A.; Woodhouse, C.A.; Lukas, J.J.; Wilkinson Kaye, M."
rwl.report.combo <- rwl.report(rwlcombo)
keymet.combo$nseries <- rwl.report.combo$nSeries
keymet.combo$FirstYear <- rwl.report.combo$firstYear
keymet.combo$LastYear <- rwl.report.combo$lastYear
keymet.combo$interrbar <- rwl.report.combo$meanInterSeriesCor
keymet.combo$interrbar.sd <- rwl.report.combo$sdInterSeriesCor
keymet.combo$ar1bar <- rwl.report.combo$meanAR1
keymet.combo$ar1bar.sd <- rwl.report.combo$sdAR1

keymet <- rbind(keymet, keymet.combo)

### Write to file ###
write.rwl(rwlcombo, fname=here("Data", "TreeRing", "Processed", "Nonhost", "DMCombo.rwl"), format="tucson")
```

## Deer Ridge

We combined ponderosa pine data collected by Veblen et al. [-@veblen2000] and subsequently recollected by Smith and Veblen in 2007.

```{r DRdat, results=F, echo=F}
## Original (Veblen et al. 2000)
rwlj <-read.rwl(veblen.series$DRI)

## Updated
rwlj1 <-read.rwl(veblen.series$DRIre)
x <- which(colnames(rwlj1) %in% colnames(rwlj))
rwlj1 <- rwlj1[,-x]

### Combine
rwlcombo <- combine.rwl(rwlj, rwlj1)
```

```{r DRfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Deer Ridge. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
rwlcombocor <- corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1)
```

```{r DRtab, tab.id="DRtab", tab.cap="Series statistics and cross-dating notes for the Deer Mountain site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlcombo)
summary.tab$rho <- round(corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab <- summary.tab %>% dplyr::select(series, first, last, rho) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation")
summary.tab$Update <- ifelse(summary.tab$Series %in% colnames(rwlj1), "Yes", "No")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r DRwrite, results=F, echo=F}
### Compile Metadata ###
keymet.combo <- keymet %>% filter(SeriesCode=="DRI") 
keymet.combo$SeriesCode <- "DRIcombo"

rwl.report.combo <- rwl.report(rwlcombo)
keymet.combo$nseries <- rwl.report.combo$nSeries
keymet.combo$FirstYear <- rwl.report.combo$firstYear
keymet.combo$LastYear <- rwl.report.combo$lastYear
keymet.combo$interrbar <- rwl.report.combo$meanInterSeriesCor
keymet.combo$interrbar.sd <- rwl.report.combo$sdInterSeriesCor
keymet.combo$ar1bar <- rwl.report.combo$meanAR1
keymet.combo$ar1bar.sd <- rwl.report.combo$sdAR1

keymet <- rbind(keymet, keymet.combo)


### Write to file ###
write.rwl(rwlcombo, fname=here("Data", "TreeRing", "Processed", "Nonhost", "DRIcombo.rwl"), format="tucson")

```

## Eldorado Canyon

We combined ponderosa pine chronologies collected by Graybill [-@graybill2002NOAAWDSPaleoclimatologya] and subsequently recollected by Smith and Veblen in 2007.

```{r ELDOdat, results=F, echo=F}
## Original - Graybill (CO534)
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", paste0("CO534", ".rwl")))

# Veblen recollection
rwlj1 <- read.rwl(veblen.series$ELDOre)
x <- which(colnames(rwlj1) %in% colnames(rwlj))
rwlj1 <- rwlj1[,-x]

### Combine
rwlcombo <- combine.rwl(rwlj, rwlj1)

```

```{r, ELDOfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Eldorado Canyon site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
x <- corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1)
```

```{r ELDOtab, tab.id="ELDOtab", tab.cap="Series statistics and cross-dating notes for the Eldorado Canyon site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlcombo)
summary.tab$rho <- round(corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab <- summary.tab %>% dplyr::select(series, first, last, rho) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation")
summary.tab$Update <- ifelse(summary.tab$Series %in% colnames(rwlj1), "Yes", "No")
summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r ELDOMeta, results=F, echo=F}
keymet.combo <- keymet %>% filter(SeriesCode=="CO534") 
keymet.combo$SeriesCode <- "ELDOcombo"
keymet.combo$Contributors <- "Graybill, D.A.; Veblen, T.T."

rwl.report.combo <- rwl.report(rwlcombo)
keymet.combo$nseries <- rwl.report.combo$nSeries
keymet.combo$FirstYear <- rwl.report.combo$firstYear
keymet.combo$LastYear <- rwl.report.combo$lastYear
keymet.combo$interrbar <- rwl.report.combo$meanInterSeriesCor
keymet.combo$interrbar.sd <- rwl.report.combo$sdInterSeriesCor
keymet.combo$ar1bar <- rwl.report.combo$meanAR1
keymet.combo$ar1bar.sd <- rwl.report.combo$sdAR1

keymet <- rbind(keymet, keymet.combo)

### Write to file
write.rwl(rwlcombo, fname=here("Data", "TreeRing", "Processed", "Nonhost", "ELDOcombo.rwl"), format="tucson")
```

## Jefferson County Colorado

We combined ponderosa pine chronologies collected by Graybill [-@graybill2002NOAAWDSPaleoclimatologyd], Woodhouse [-@woodhouse2006NOAAWDSPaleoclimatologyg] and subsequently recollected by Smith and Veblen in 2007.

```{r JCdat, results=F, echo=F}
## Original - Graybill CO539
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO539.rwl"))

## Updated - Woodhouse CO608
rwlj1 <-read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO608.rwl"))
x <- which(colnames(rwlj1) %in% colnames(rwlj))
if(length(x)>0){
  rwlj1 <- rwlj1[,-x]
}

rwlcombo <- combine.rwl(rwlj, rwlj1)

## Updated - Veblen JEFFre
rwlj2 <-read.rwl(veblen.series$JEFFre)
x <- which(colnames(rwlj2) %in% colnames(rwlcombo))
if(length(x)>0){
  rwlj2 <- rwlj2[,-x]
}

rwlcombo <- combine.rwl(rwlcombo, rwlj2)
```

```{r JCfig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Jefferson County Open Space site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
x <- corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1)
```

```{r JCtab, tab.id="JCtab", tab.cap="Series statistics and cross-dating notes for the Jefferson County Open Space site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlcombo)
summary.tab$rho <- round(corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab <- summary.tab %>% dplyr::select(series, first, last, rho) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation")
summary.tab$Update <- ifelse(summary.tab$Series %in% c(colnames(rwlj1),colnames(rwlj2)), "Yes", "No")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r JCwrite, results=F, echo=F}
```

```{r JEFFMeta, results=F, echo=F}
rwlcombo <- rwlcombo[,!(colnames(rwlj) ==("J2_8"))] # remove series with poorly correlated data

keymet.combo <- keymet %>% filter(SeriesCode=="CO539") 
keymet.combo$SeriesCode <- "JEFFcombo"
keymet.combo$Contributors <- "Graybill, D.A.; Woodhouse, C.A.; Lukas, J.J.; Nepstad-Thornberry, C.; Veblen, T.T."

rwl.report.combo <- rwl.report(rwlcombo)
keymet.combo$nseries <- rwl.report.combo$nSeries
keymet.combo$FirstYear <- rwl.report.combo$firstYear
keymet.combo$LastYear <- rwl.report.combo$lastYear
keymet.combo$interrbar <- rwl.report.combo$meanInterSeriesCor
keymet.combo$interrbar.sd <- rwl.report.combo$sdInterSeriesCor
keymet.combo$ar1bar <- rwl.report.combo$meanAR1
keymet.combo$ar1bar.sd <- rwl.report.combo$sdAR1

keymet <- rbind(keymet, keymet.combo)

### Write to file
write.rwl(rwlcombo, fname=here("Data", "TreeRing", "Processed", "Nonhost", "JEFFcombo.rwl"), format="tucson")
```

## Kassler

We combined ponderosa pine chronologies collected by Fritts [-@fritts2002NOAAWDSPaleoclimatology] and Graybill [-@graybill2002NOAAWDSPaleoclimatologyb].

```{r Kasslerdat, results=F, echo=F}
# Original - Fritts CO030
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO030.rwl"))

## Updated - Graybill CO541
filej <- list.files(, full.names=T, ignore.case=T)
rwlj1 <-read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa","CO541.rwl"))
x <- which(colnames(rwlj1) %in% colnames(rwlj))
if(length(x)>0){
  rwlj1 <- rwlj1[,-x]
}

### Combine
rwlcombo <- combine.rwl(rwlj, rwlj1)
```

```{r KasslerFig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Kassler site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
x <- corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1)
```

```{r KasslerTab, tab.id="Kasslertab", tab.cap="Series statistics and cross-dating notes for the Kassler site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlcombo)
summary.tab$rho <- round(corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab <- summary.tab %>% dplyr::select(series, first, last, rho) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation")
summary.tab$Update <- ifelse(summary.tab$Series %in% c(colnames(rwlj1),colnames(rwlj2)), "Yes", "No")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r KaslerMeta, results=F, echo=F}
keymet.combo <- keymet %>% filter(SeriesCode=="CO030") 
keymet.combo$SeriesCode <- "KasslerCombo"
keymet.combo$Contributors <- "Fritts, H.C.; Holmes, R.L.; Graybill, D.A."

rwl.report.combo <- rwl.report(rwlcombo)
keymet.combo$nseries <- rwl.report.combo$nSeries
keymet.combo$FirstYear <- rwl.report.combo$firstYear
keymet.combo$LastYear <- rwl.report.combo$lastYear
keymet.combo$interrbar <- rwl.report.combo$meanInterSeriesCor
keymet.combo$interrbar.sd <- rwl.report.combo$sdInterSeriesCor
keymet.combo$ar1bar <- rwl.report.combo$meanAR1
keymet.combo$ar1bar.sd <- rwl.report.combo$sdAR1

keymet <- rbind(keymet, keymet.combo)

### Write to file
write.rwl(rwlcombo, fname=here("Data", "TreeRing", "Processed", "Nonhost", "KaslerCombo.rwl"), format="tucson")
```

## South Fork

We combined ponderosa pine chronologies collected by Woodhouse et al. [-@woodhouse2006NOAAWDSPaleoclimatologyh] and Woodhouse and Losleben [-@woodhouse2016NOAAWDSPaleoclimatology].

```{r SFdat, results=F, echo=F}
## Original --> CO625 
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO625.rwl"))

## Udpdated --> CO653
rwlj1 <- read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO653.rwl"))
x <- which(colnames(rwlj1) %in% colnames(rwlj))
if(length(x)>0){
  rwlj1 <- rwlj1[,-x]
}

### Combine
rwlcombo <- combine.rwl(rwlj, rwlj1)
```

```{r SFFig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the South Fork site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
x <- corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1)
```

```{r SFTab, tab.id="SFtab", tab.cap="Series statistics and cross-dating notes for the South Fork site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlcombo)
summary.tab$rho <- round(corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab <- summary.tab %>% dplyr::select(series, first, last, rho) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation")
summary.tab$Update <- ifelse(summary.tab$Series %in% c(colnames(rwlj1),colnames(rwlj2)), "Yes", "No")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r SouthForkMeta, results=F, echo=F}
keymet.combo <- keymet %>% filter(SeriesCode=="CO625") 
keymet.combo$SeriesCode <- "SFcombo"

rwl.report.combo <- rwl.report(rwlcombo)
keymet.combo$nseries <- rwl.report.combo$nSeries
keymet.combo$FirstYear <- rwl.report.combo$firstYear
keymet.combo$LastYear <- rwl.report.combo$lastYear
keymet.combo$interrbar <- rwl.report.combo$meanInterSeriesCor
keymet.combo$interrbar.sd <- rwl.report.combo$sdInterSeriesCor
keymet.combo$ar1bar <- rwl.report.combo$meanAR1
keymet.combo$ar1bar.sd <- rwl.report.combo$sdAR1

keymet <- rbind(keymet, keymet.combo)

### Write to file
write.rwl(rwlcombo, fname=here("Data", "TreeRing", "Processed", "Nonhost", "SFcombo.rwl"), format="tucson")
```

## Turkey Creek Bluff

We combined ponderosa pine chronologies collected by Graybill [-@graybill2002NOAAWDSPaleoclimatologyc] and Woodhouse et. al [-@woodhouse2019NOAAWDSPaleoclimatologya].

```{r TCdat, results=F, echo=F}
## CO548
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO548.rwl"))

## CO675
rwlj1 <- read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO675.rwl"))
x <- which(colnames(rwlj1) %in% colnames(rwlj))
if(length(x)>0){
  rwlj1 <- rwlj1[,-x]
}

### Combine
rwlcombo <- combine.rwl(rwlj, rwlj1)
```

```{r TCFig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Turkey Creek Bluff site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
x <- corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1)
```

```{r TCTab, tab.id="TCtab", tab.cap="Series statistics and cross-dating notes for the Turkey Creek Bluff site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlcombo)
summary.tab$rho <- round(corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab <- summary.tab %>% dplyr::select(series, first, last, rho) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation")
summary.tab$Update <- ifelse(summary.tab$Series %in% c(colnames(rwlj1),colnames(rwlj2)), "Yes", "No")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r TCMeta, results=F, echo=F}
keymet.combo <- keymet %>% filter(SeriesCode=="CO548") 
keymet.combo$SeriesCode <- "TCcombo"
keymet.combo$Contributors <- "Graybill, D.A.; Woodhouse, C.A.; Losleben, M.V.; Chowanski, K.; Lukas, J.J"

rwl.report.combo <- rwl.report(rwlcombo)
keymet.combo$nseries <- rwl.report.combo$nSeries
keymet.combo$FirstYear <- rwl.report.combo$firstYear
keymet.combo$LastYear <- rwl.report.combo$lastYear
keymet.combo$interrbar <- rwl.report.combo$meanInterSeriesCor
keymet.combo$interrbar.sd <- rwl.report.combo$sdInterSeriesCor
keymet.combo$ar1bar <- rwl.report.combo$meanAR1
keymet.combo$ar1bar.sd <- rwl.report.combo$sdAR1

keymet <- rbind(keymet, keymet.combo)

### Write to file
write.rwl(rwlcombo, fname=here("Data", "TreeRing", "Processed", "Nonhost", "TCcombo.rwl"), format="tucson")
```

## Van Bibber Creek

We combined ponderosa pine chronologies collected by Graybill [-@graybill1997NOAAWDSPaleoclimatology], Woodhouse et al. [-@woodhouse2010NOAAWDSPaleoclimatologya] and subsequently recollected by Smith and Veblen in 2007.

```{r VANBdat, results=F, echo=F}
## Original - CO550
rwlj <-read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO550.rwl"))

## Updated - CO641
rwlj1 <- read.rwl(here("Data", "TreeRing", "Raw", "Nonhost", "ITRDB", "data", "pub", "data", "paleo", "treering", "measurements", "northamerica", "usa", "CO641.rwl"))
x <- which(colnames(rwlj1) %in% colnames(rwlj))
if(length(x)>0){
  rwlj1 <- rwlj1[,-x]
}
rwlcombo <- combine.rwl(rwlj, rwlj1)

## Veblen Updated
rwlj2 <- read.rwl(veblen.series$VANBre)
x <- which(colnames(rwlj2) %in% colnames(rwlcombo))
if(length(x)>0){
  rwlj2 <- rwlj2[,-x]
}

rwlcombo <- combine.rwl(rwlcombo, rwlj2)
```

```{r VANBFig, fig.width=4.5, fig.height=4.5, fig.cap="The correlations between each tree-ring series and a main chronology built from all other series from the Van Bibber Creek site. Correlations were calculated in 50-year segments that overlap by 25 years. Blue segments show period that correlate well with the main chronology (p<0.1). Red segments show poor correlation with the main chronology (p>0.1). Green segments show periods do not overlap with the main chronology."}
x <- corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1)
```

```{r VANBTab, tab.id="TCtab", tab.cap="Series statistics and cross-dating notes for the Van Bibber Creek site. Spearman's correlations were calculated over the common period of overlap between the series and a main chronology built from all other series."}
summary.tab <- summary.rwl(rwlcombo)
summary.tab$rho <- round(corr.rwl.seg(rwlcombo, seg.length=50, pcrit=0.1, make.plot=F)$overall[,1],2)
summary.tab <- summary.tab %>% dplyr::select(series, first, last, rho) %>% mutate(first=as.character(first), last=as.character(last))
colnames(summary.tab) <- c("Series", "First\nyear", "Last\nyear", "Spearman's correlation")
summary.tab$Update <- ifelse(summary.tab$Series %in% c(colnames(rwlj1),colnames(rwlj2)), "Yes", "No")

summary.tab[order(summary.tab$Series),] %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)
```

```{r VANBMeta, results=F, echo=F}
keymet.combo <- keymet %>% filter(SeriesCode=="CO550") 
keymet.combo$SeriesCode <- "VANBcombo"
keymet.combo$Contributors <- "Graybill, D.A.; 	
Woodhouse, C.A.; Lukas, J.J.; Veblen. T.T."

rwl.report.combo <- rwl.report(rwlcombo)
keymet.combo$nseries <- rwl.report.combo$nSeries
keymet.combo$FirstYear <- rwl.report.combo$firstYear
keymet.combo$LastYear <- rwl.report.combo$lastYear
keymet.combo$interrbar <- rwl.report.combo$meanInterSeriesCor
keymet.combo$interrbar.sd <- rwl.report.combo$sdInterSeriesCor
keymet.combo$ar1bar <- rwl.report.combo$meanAR1
keymet.combo$ar1bar.sd <- rwl.report.combo$sdAR1

keymet <- rbind(keymet, keymet.combo)

### Write to file
write.rwl(rwlcombo, fname=here("Data", "TreeRing", "Processed", "Nonhost", "VANBcombo.rwl"), format="tucson")
```

```{r writeMetadata, results=F, echo=F}
write.csv(keymet, here("Data", "TreeRing", "Processed", "Nonhost", "Nonhost-Metadata.csv"), row.names=F)
```

# References
