---
title: "Identify control chronologies for detecting defoliation by the western spruce budworm in Douglas-fir radial growth time series based on climate and tree-ring data for: Drought may initiate western spruce budworm outbreaks, but multi-year periods of increased moisture availability promote widespread defoliation"
author: "Sarah J. Hart, Olivia Santiago, Josh Carrell, and Thomas T. Veblen"
email: "sarah.hart@colostate.edu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
   officedown::rdocx_document:
     reference_docx: Template.docx
bibliography: references.bib
csl: "`r here::here('citationstyle.csl')`"
link-citations: true
urlcolor: blue
linkcolor: blue
citationcolor: blue
---

```{r setup,include=FALSE, results='hide'}
knitr::opts_chunk$set(
  echo = T,
	message = FALSE,
	warning = FALSE,
	progress = FALSE,
	cache = FALSE,
	dpi = 300,
  fig.align = 'center'
)

# Set seed
set.seed(2020)

### Import libraries
if (!require("pacman")){install.packages("pacman")}
pacman::p_load(
  here, # easy file structure
  tidyverse, # data manipulation
  ggplot2, # figures
  patchwork, # combining figures
  dplR, # basic dendrochronology
  zoo, # time series
  knitr, # markdown documents
  kableExtra, # tables in markdown
  flextable, # other table options
  officer,
  bookdown,
  sf, # spatial data (new pkg)
  ncdf4, #import ncdf files
  terra,  # newer package for working with raster data
  raster  # older package for working with raster data (code still uses old function for creating raster from array) 
) 
proj.proj <- 4629

# Set directory structure for project
dir.create(here("Data", "Spatial"), showWarnings = FALSE)
dir.create(here("Results"), showWarnings = FALSE)
```

# Overview
Briefly, here we matched nonhost sites with host sites based the highest Pearson's correlation, subject to the following constraints:

1.  the nonhost chronology overlapped the entire host chronology

2.  the nonhost chronology was less than 125 km away from the host site.

Finally, we removed any sites where the host chronology was poorly correlated (r\<0.2) with either the SC-PDSI or the available nonhost chronologies. Results of this matching are presented in Table \@ref(tab:hostxnonhost)).

```{r TreeRingData, results='hide'}
host.meta <- read.csv(here("Data", "TreeRing", "Processed", "Host", "host-metadata.csv"))
nonhost.meta <- read.csv(here("Data", "TreeRing", "Processed", "Nonhost", "nonhost-metadata.csv")) %>% filter(is.na(Lat)==F)

nonhost.sites <- st_as_sf(x = nonhost.meta,  coords = c("Lon", "Lat"), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>% st_transform(proj.proj)

host.sites <- st_as_sf(x = host.meta, coords = c("Lon", "Lat"), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>% st_transform(proj.proj)
```

```{r NADA, results='hide'}
# Read in NADA time series
nada.hosts <- read.csv(here("Data", "nada-hosts.csv"))
```

```{r HostXNonhost, results='hide'}
### READ IN RINGWIDTH DATA
# Host data
host.rwls <- list.files(here("Data", "TreeRing", "Processed", "Host"), pattern=".rwl", full.names=T)
names(host.rwls) <- gsub(".rwl", "", unlist(list.files(here("Data", "TreeRing", "Processed", "Host"), pattern=".rwl", full.names=F)))
host.rwls <- lapply(host.rwls, read.rwl)

# Nonhost data
nonhost.rwls <- list.files(here("Data", "TreeRing", "Processed", "NonHost"), full.names = T, pattern=".rwl")
names(nonhost.rwls) <- gsub(".rwl", "", unlist(list.files(here("Data", "TreeRing", "Processed", "NonHost"), pattern=".rwl", full.names=F)))
nonhost.rwls <- lapply(nonhost.rwls, read.rwl)
namez <- names(nonhost.rwls) 
bad.namez <- which(namez %in% c("CO532", "CO598", "DRI", "DRIre", "CO534", "ELDOre", "CO539", "CO608", "JEFFre", "CO030", "CO541", "CO625", "CO653", "CO548", "CO675", "CO550", "CO641", "VANBre"))
nonhost.rwls <- nonhost.rwls[-bad.namez]

### DETREND ###
# Host data
host.rwis <- lapply(host.rwls, detrend, method = "ModNegExp")
host.rwis <- lapply(host.rwis, detrend, method = "Spline", nyrs = 30)

# Nonhost data
nonhost.rwis <- lapply(nonhost.rwls, detrend, method = "ModNegExp")
nonhost.rwis <- lapply(nonhost.rwis, detrend, method = "Spline", nyrs = 30)

### BUILD CHRONOLOGIES
# Host data
host.crns <- lapply(host.rwis, chron, prewhiten = TRUE, prefix="")
host.crns <- lapply(host.crns, "[",c(-1, -3))
host.crns.ts <- lapply(host.crns, FUN=function(x){return( ts(x, start=min(as.numeric(row.names(x)))) )})
host.crns.mts<- do.call(cbind, host.crns.ts)

# Nonhost data
nonhost.crns <- lapply(nonhost.rwis, chron, prewhiten = TRUE, prefix="")
nonhost.crns <- lapply(nonhost.crns, "[",c(-1, -3))
nonhost.crns.ts <- lapply(nonhost.crns, FUN=function(x){return( ts(x, start=min(as.numeric(row.names(x)))))})
nonhost.crns.mts<- do.call(cbind, nonhost.crns.ts)

## CALCULATE DISTANCES
host.pts <-  st_as_sf(x = host.meta, coords = c("Lon", "Lat"), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>%  st_transform("epsg:26913")

nonhost.pts <-  st_as_sf(x = nonhost.meta[!is.na(nonhost.meta$Lat)==T,], coords = c("Lon", "Lat"), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>% st_transform("epsg:26913") %>%  filter(SeriesCode %in% names(nonhost.crns))

dists <- st_distance(host.pts, nonhost.pts, by_element=F) 
row.names(dists)<- host.pts$SeriesCode
colnames(dists)<- nonhost.pts$SeriesCode

### CALCUALTE CORRELATIONS
cor.mat <- cor(cbind(host.crns.mts, nonhost.crns.mts), use="pairwise.complete.obs")
row.names(cor.mat) <- c(colnames(host.crns.mts), colnames(nonhost.crns.mts))
colnames(cor.mat) <- c(colnames(host.crns.mts), colnames(nonhost.crns.mts))

# Summarize
res <- data.frame(host=host.meta$SeriesCode, nonhost1=NA, nonhost2=NA, nonhost3=NA,  nonhost1.r=NA, nonhost2.r=NA, nonhost3.r=NA, nonhost1.d=NA, nonhost2.d=NA, nonhost3.d=NA, r.hostxNADA=NA) # create data frame to hold results
for(j in res$host){
  
  # (1) limit to chronologies that overlap the entire host record
  lastyrj <- host.meta %>% dplyr::filter(SeriesCode == j) %>% pull(LastYear)
  firstyrj <- host.meta %>% dplyr::filter(SeriesCode == j) %>% pull(FirstYear)
  firstyrj <- ifelse(firstyrj<1650, 1650, firstyrj)
  nonhost.long <- nonhost.meta %>% filter(SeriesCode %in% nonhost.pts$SeriesCode & LastYear>=lastyrj & FirstYear<=firstyrj)  %>% pull(SeriesCode)
  
  # (2) limit to sites within 100 km 
  dists.j <- as.data.frame(dists)
  dists.j <- dists[j, nonhost.long]
  dists.j <- unclass(dists.j)
  dists.j <- dists.j[dists.j <= 125000] 
  
  # (3) determine top 3 correlations
  cor.matj <- cor.mat[j, names(dists.j)]
  top3 <- sort(cor.mat[j, names(dists.j)],decreasing=T)[1:3]
  res[res$host==j, c("nonhost1", "nonhost2", "nonhost3")] <- names(top3)
  res[res$host==j, c("nonhost1.r", "nonhost2.r", "nonhost3.r")] <- cor.matj[names(top3)]
  res[res$host==j, c("nonhost1.d", "nonhost2.d", "nonhost3.d")] <-  dists.j[names(top3)]
  
  # (4) correlation between host and NADA
  nada.ts <- ts(nada.hosts[,j], start=min(nada.hosts$X))
  res[res$host==j, ]$r.hostxNADA <- cor(cbind(host.crns.mts[,colnames(host.crns.mts)==j], nada.ts), use="pairwise.complete.obs")[1,2]

}
res$mean.r <- res %>% dplyr::select(nonhost1.r:nonhost3.r) %>% rowMeans(na.rm=T)
res$notes <- NA
res2 <- res

res2 <- res2 %>% filter(r.hostxNADA>=0.2) # Estes park, Frankenberger Point, or South Hollowell do not correlate well with PDSI
res <- res %>% mutate(notes=ifelse(r.hostxNADA<0.2, "Removed due to poor correlation with PDSI", NA))

write.csv(res, here("Results", "hostXnonhost.csv"), row.names=F)
write.csv(res2, here("Results", "hostXnonhost-subset.csv"), row.names=F)
```

```{r summary,  tab.cap="Summary of nonhost chronologies matched with each host site. For the top three nonhost sites, the Pearson's correlation coefficient (r) and distance (d) between the host and nonhost sits are displayed in parentheses below the name of each series.", tab.id='hostxnonhost'}

res <- read.csv(here("Results", "hostXnonhost.csv"))
res <- res %>%  mutate(nonhost1.r=round(nonhost1.r,digits=2), nonhost2.r=round(nonhost2.r,digits=2), nonhost3.r=round(nonhost3.r,digits=2), nonhost1.d=round(nonhost1.d/1000,digits=0), nonhost2.d=round(nonhost2.d/1000,digits=0), nonhost3.d=round(nonhost3.d/1000,digits=0), r.hostxNADA=round(r.hostxNADA,digits=2))

res <- res %>% unite("Control1", nonhost1, nonhost1.r, remove = TRUE, sep=" (r=") %>% mutate(Control1=paste0(Control1, "; ")) %>% unite("Control1", Control1, nonhost1.d, remove = TRUE, sep="d=") %>% mutate(Control1=paste0(Control1, " km)")) 

res <- res %>% unite("Control2", nonhost2, nonhost2.r, remove = TRUE, sep=" (r=") %>% mutate(Control2=paste0(Control2, "; ")) %>% unite("Control2", Control2, nonhost2.d, remove = TRUE, sep="d=") %>% mutate(Control2=paste0(Control2, " km)")) 

res <- res %>% unite("Control3", nonhost3, nonhost3.r, remove = TRUE, sep=" (r=") %>% mutate(Control3=paste0(Control3, "; ")) %>% unite("Control3", Control3, nonhost3.d, remove = TRUE, sep="d=") %>% mutate(Control3=paste0(Control3, " km)")) 
  
res <- res %>% dplyr::select(host, Control1, Control2, Control3, r.hostxNADA, notes)
colnames(res) <- c("Host site", "Nonhost control site 1", "Nonhost control site 2", "Nonhost control site 3", "Correlation with PDSI", "Notes")

res %>% flextable() %>% flextable::align(j=-1, align = "center", part = "all") %>% set_table_properties(layout = "autofit", width=1)

```
